---
title: "How is Hate Crime Rate Affected by Economic and Ethnic Demographics"
author: "Group D: Audrey Bertin and Marissa Freed"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
library(readr)
library(GGally)

hate_crimes <- read_csv("~/Desktop/MTH 220/group-D-data.csv")
```

```{r}
hate_crimes_minusDC <- hate_crimes %>% 
  filter(state != "District of Columbia")
```


# Abstract

This report considers the economic and ethnic demographics of states and how they are connected with states' hate crime rates. In order to determine which characteristics were most strongly correlated with hate crimes, several variables including Gini Index, electoral results (2016 Presidential), and the share of the population with a high school degree were compared with hate crime rates as measured by the Federal Bureau of Investigation (FBI) and Southern Poverty Law Center (SPLC). It was determined that few variables had a very strong correlation, and the variables among both hate crime rates that were the most correlated were associated with the proportion of Trump voters, median income, and inequality. Multiple regression models were also created for each hate crime rate source, though collinearity among some of the variables likely affected their coefficients.

# Introduction

Mentions of hate crimes in the news and through social media have appeared in larger numbers recently than in the past, with a distinct spike in incidents since Donald Trump's election in November 2016. We wanted to see whether Trump voter rate was associated with hate crimes. We were also interested in what factors determine hate crime rate, and whether certain demographic statistics affect hate crimes more than others, in order to better learn how to predict them. Several specific questions of interest include:

*1) Are Trump voter rate and share of population with a high school degree the two variables most strongly associated with hate crimes? Do the directions of their relationships match what is expected?*

Due to the increase in hate crimes after the 2016 election, and the fact that on average, Trump voters had a lower level of education than Clinton voters ([source](http://fivethirtyeight.com/features/education-not-income-predicted-who-would-vote-for-trump/)), we study was particularly interested in testing the hypothesis that `share_voters_voted_Trump` and `share_population_with_high_school_degree` are both more strongly correlated with hate crime rate than the other variables. Additionally, based on background information, we wanted to determine whether their directions matched what was expected: a positive correlation for Trump voter rate and a negative correlation for share of the population with a high school degree. 


*2) What model can we create to best predict hate crime rate?*

This information would be important in order to help law enforcement agencies determine where special lookout for hate crimes should be used, and could also be used by people who are looking to move to a particular area. If a person is in a group that is affected by hate crimes (LGBT, Jewish, etc.), then that person could look at the demographics for the place they are interested in moving to and determine the danger level of it. 

Additionally, this data would be helpful in identifying certain factors wthin states (and the same model could be used within states for identifying communities) that can be targeted (such as economic inequality, uneducation) to improve in order to lower the hate crime rate.


# Data

The dataset comes from FiveThirtyEight and is a combination of hate crime data from the Federal Bureau of Investigation (FBI) and Southern Poverty Law Center (SPLC) and data on economic and ethnic demographics collected by the US Census Bureau. The FBI hate crime rate is an average of the hate crime rates from 2010 to 2015, and the SPLC hate crime rate is from 2016.

All of the variables except `state` are continuous and numerical:

1) 'median_household_income' represents the average household income in the state as measured in 2016 (dollars).

2) 'share_unemployed_seasonal' represents the share (percentage) of the population that was unemployed in September 2016, adjusted for the season of the year.

3) 'share_population_in_metro_areas' represents the proportion (percentage) of the population that lives in metropolitan/urban areas.

4) 'share_population_with_high_school_degree' is the proportion (percentage) of adults aged 25+ who graduated from high school.

5) 'share_non_citizen' is the proportion (percentage) of the population that are not U.S. citizens.

6) 'share_white_poverty' is the proportion (percentage) of white residents who are living in poverty in the area.

7) 'gini_index' is the 2015 gini index value. The gini value is a measure of economic inequality between 0 (0%) and 1 (100%) with 0 representing perfect equality and 1 representing perfect inequality)

8) 'share_non_white' is the proportion (percentage) of the population that was not white in 2015

9) 'share_voters_voted_trump' is the proportion (percentage) of voters who voted for Donald Trump in the 2016 election.

The response variables `hate_crimes_per_100k_splc` and `avg_hatecrimes_per_100k_fbi` are both measured as the number of hate crimes per 100,000 people.

The sample, each state's summary, is taken from the population of the entire United States. Most explanatory variables were calculated by summarizing Census Bureau data collected for each state, while `share_voters_voted_trump` was calculated with 2016 election data. These values likely do not account for all people in the state since some people are not legal residents and therefore would not fill out the Census. 

In order to calculate hate crime rates, the FBI bases its rate on official criminal charges for hate crime, and does not provide a way for individuals to submit incident reports.

The SPLC hate crime rate is based on a collection of incident reports submitted on [their website](https://www.splcenter.org/reporthate). As a result, this self-reported rate is much lower than the FBI rate for all states.

It is important to note that both reported hate crime rates are likely to be inaccurate when compared to the true rate. This is due to the fact that many hate crimes are not reported, either by the victim to local law enforcement or from police to the FBI. In order for the FBI to receive hate crime reports, local law enforcement agencies must actively collect and submit this data to the state, which then submits it to the FBI. This often does not happen. Thousands of police agencies have also completely opted out of the FBI's hate crime program, and don't report any data. 

Pro-Publica did an [investigation](https://www.propublica.org/article/why-america-fails-at-gathering-hate-crime-statistics) into this problem, and found that while only 6,212 hate crimes were reported to the FBI in 2016, true estimates place the number at 250,000 hate crimes. This means that the hate crime data available is likely quite inaccurate.

Each state's hate crime rate accuracy is also likely to be different, as states define hate crimes differently from one another--for example, Alabama's hate-crime statute does not apply to crimes based on sexual orientation, while Connecticut's statute includes crimes based on sexual orientation as well as gender identity. As a result, many cases in that may be considered hate crimes in some states are not reported as such in others.

Since the data are all numerical, and there are many variables recorded, we chose to use multiple regression to answer our research questions, adjusting for the number of variables by looking at R-squared-adj.

# Methods

In order to answer both research questions, we performed one set of calculations using the FBI response variable and another set with the SPLC response variable.

In each set of calculations, to answer the first research question, we paired each explanatory variable with the hate crime rate and determined the direction and correlation coefficient. We then identified the three explanatory variables with the strongest correlations, and determined:

  1) Whether `share_voters_voted_trump` and `share_population_with_high_school_degree` were the top two most correlated variables (in either order).
  
  2) If they were, did the direction of their relationship match what we expected?

While we did not remove any outliers for the official results, we also computed a second round of calculations, the same way as described previously, on the dataset with the outlier (Washington, D.C.) removed. This was done in order to see what effect the outlier was having on our results.

To answer the second research question, we performed backward elimination using p-values to determine the best multiple regression model for each data source.

Finally, we reported whether the models we created to answer the second question met the conditions for inference.


# Results
### (Includes Basic Descriptive Statistics)

In order determine which variables were most strongly correlated with hate crime rate, a regression line for each explanatory variable was calculated. This was completed twice: once for the FBI response variable, and once for the SPLC response variable. 

During these calculations, it became clear that the FBI hate crime rate for Washington D.C. was an outlier, at more than twice the size of the next largest rate. This is likely due to the fact that Washington, D.C. is a city, while all of the other cases are states. While not used in the final conclusions, the regression lines for each variable were recalculated with the data for Washington D.C. removed, in order to determine if the most strongly correlated explanatory variables were different.

The SPLC hate crime rate for Washington, D.C. also appeared to be an outlier, so the same calculations without the district included were done just as with the FBI data.

The results for the calculations are below:

## Using FBI Data:

### 1) Median household income: (R-squared = 0.1013)


```{r, warning = FALSE}
income_FBI <- lm(avg_hatecrimes_per_100k_fbi ~ median_household_income, data = hate_crimes)
summary(income_FBI)


ggplot(data = hate_crimes, aes(x = median_household_income, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.05863)

```{r}
income_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ median_household_income, data = hate_crimes_minusDC)
summary(income_FBI_noDC)
```



### 2) Unemployment rate (seasonally adjusted): (R-squared = 0.005467)

```{r, warning = FALSE}
unemployment_FBI <- lm(avg_hatecrimes_per_100k_fbi ~ share_unemployed_seasonal, data = hate_crimes)
summary(unemployment_FBI)

ggplot(data = hate_crimes, aes(x = share_unemployed_seasonal, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.01998)

```{r}
unemployment_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ share_unemployed_seasonal, data = hate_crimes_minusDC)
summary(unemployment_FBI_noDC)
```

### 3) Share of the population living in a metro area: (R-squared = 0.04403)

```{r, warning = FALSE}
metro_FBI <- lm(avg_hatecrimes_per_100k_fbi ~ share_population_in_metro_areas, data = hate_crimes)
summary(metro_FBI)

ggplot(data = hate_crimes, aes(x = share_population_in_metro_areas, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.009974)

```{r}
metro_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ share_population_in_metro_areas, data = hate_crimes_minusDC)
summary(metro_FBI_noDC)
```

### 4) Share of the population with a high school degree: (R-squared = 0.02694)

```{r, warning = FALSE}
high_school_FBI <-lm(avg_hatecrimes_per_100k_fbi ~ share_population_with_high_school_degree, data = hate_crimes)
summary(high_school_FBI)

ggplot(data = hate_crimes, aes(x = share_population_with_high_school_degree, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.05112)

```{r}
high_school_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ share_population_with_high_school_degree, data = hate_crimes_minusDC)
summary(high_school_FBI_noDC)
```

### 5) Share of non-citizens: (R-squared = 0.09054)

```{r, warning = FALSE}
noncitizens_FBI <-lm(avg_hatecrimes_per_100k_fbi ~ share_non_citizen, data = hate_crimes)
summary(noncitizens_FBI)

ggplot(data = hate_crimes, aes(x = share_non_citizen, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```


Minus D.C. (R-squared = 0.02562)

```{r}
noncitizens_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ share_non_citizen, data = hate_crimes_minusDC)
summary(noncitizens_FBI_noDC)
```

### 6) Share of white poverty: (R-squared = 0.06765)

```{r, warning = FALSE}
white_poverty_FBI <- lm(avg_hatecrimes_per_100k_fbi ~ share_white_poverty, data = hate_crimes)
summary(white_poverty_FBI)

ggplot(data = hate_crimes, aes(x = share_white_poverty, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.003741)

```{r}
white_poverty_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ share_white_poverty, data = hate_crimes_minusDC)
summary(white_poverty_FBI_noDC)
```

### 7) Gini index (measure of inequality): (R-squared = 0.1775)

```{r, warning = FALSE}
gini_FBI <- lm(avg_hatecrimes_per_100k_fbi ~ gini_index, data = hate_crimes)
summary(gini_FBI)

ggplot(data = hate_crimes, aes(x = gini_index, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.003151)

```{r}
gini_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ gini_index, data = hate_crimes_minusDC)
summary(gini_FBI_noDC)
```

### 8) Share of residents who are non-white: (R-squared = 0.006976)

```{r, warning = FALSE}
nonwhite_FBI <- lm(avg_hatecrimes_per_100k_fbi ~ share_non_white, data = hate_crimes)
summary(nonwhite_FBI)

ggplot(data = hate_crimes, aes(x = share_non_white, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```


Minus D.C. (R-squared = 0.04613)

```{r}
nonwhite_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ share_non_white, data = hate_crimes_minusDC)
summary(nonwhite_FBI_noDC)
```

### 9) Share of voters who voted for Trump in 2016: (R-squared = 0.2546)

```{r, warning = FALSE}
trump_FBI <- lm(avg_hatecrimes_per_100k_fbi ~ share_voters_voted_trump, data = hate_crimes)
summary(trump_FBI)

ggplot(data = hate_crimes, aes(x = share_voters_voted_trump, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.02999)

```{r}
trump_FBI_noDC <- lm(avg_hatecrimes_per_100k_fbi ~ share_voters_voted_trump, data = hate_crimes_minusDC)
summary(trump_FBI_noDC)
```


For the FBI data, the explanatory variables with the three highest correlations were: 

1) `share_voters_voted_trump` (R-squared = 0.2546, r = -0.504)

2) `gini_index` (R-squared = 0.1775, r = 0.421)

3) `median_household_income` (R-squared = 0.1013, r = 0.318)

Of the two variables that we predicted would be the most correlated, only one appears in this list. Additionally, the direction of the relationship does not match what was expected. `share_voters_voted_trump` has a negative relationship with hate crime rate here.


When the Washington, D.C. outlier was removed, the three most correlated variables were:

1) `median_household_income` (R-squared = 0.05863, r = 0.242)

2) `share_population_with_high_school_degree` (R-squared = 0.05112, r = 0.226)

3) `share_non_white` (R-squared = 0.04613, r = 0.215)

In this list, the second variable that we predicted would be most correlated appears, though `share_voters_voted_trump` is no longer included. The direction of the relationship between `share_population_with_high_school_degree` also does not match what we expected--it is positive rather than negative.

The correlated variables with and without Washington D.C. are relatively different, with only median household income appearing in both lists.


### Fitting the best model

In order to fit the best model, backward-elimination was used. The best fit model was calculated while including the data for Washington, D.C. Below is the model with all variables included.

```{r, warning = FALSE}

FBI_model_all <- lm(avg_hatecrimes_per_100k_fbi ~ median_household_income + share_unemployed_seasonal + share_population_in_metro_areas + share_population_with_high_school_degree + share_non_citizen + share_white_poverty + gini_index + share_non_white + share_voters_voted_trump, data = hate_crimes)
summary(FBI_model_all)

```


The best model included only two variables that were statistically significant, `share_population_with_high_school_degree` (p-value = 3.29e-05) and `gini_index` (p-value = 5.55e-07)

```{r, warning = FALSE}
FBI_model_best <- lm(avg_hatecrimes_per_100k_fbi ~ share_population_with_high_school_degree + gini_index,  data = hate_crimes)
summary(FBI_model_best)
```

This model has the equation:

avg_hatecrimes_per_100k_fbi = (31.287 x share_population_with_high_school_degree) + (64.324 x gini_index) - 54.017 


## Using SPLC Data

The same calculations as described above were also completed for the SPLC data.


### 1) Median household income: (R-squared = 0.123)

```{r, warning = FALSE}
income_SPLC <- lm(hate_crimes_per_100k_splc ~ median_household_income, data = hate_crimes)
summary(income_SPLC)

ggplot(data = hate_crimes, aes(x = median_household_income, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```


Minus D.C. (R-squared = 0.08232)

```{r}
income_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ median_household_income, data = hate_crimes_minusDC)
summary(income_SPLC_noDC)
```


### 2) Unemployment rate (seasonally adjusted): (R-squared = 0.006938)

```{r, warning = FALSE}
unemployment_SPLC <- lm(hate_crimes_per_100k_splc ~ share_unemployed_seasonal, data = hate_crimes)
summary(unemployment_SPLC)

ggplot(data = hate_crimes, aes(x = share_unemployed_seasonal, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.01729)

```{r}
unemployment_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ share_unemployed_seasonal, data = hate_crimes_minusDC)
summary(unemployment_SPLC_noDC)
```

### 3) Share of the population living in a metro area: (R-squared = 0.02631)

```{r, warning = FALSE}
metro_SPLC <- lm(hate_crimes_per_100k_splc ~ share_population_in_metro_areas, data = hate_crimes)
summary(metro_SPLC)

ggplot(data = hate_crimes, aes(x = share_population_in_metro_areas, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.005875)

```{r}
metro_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ share_population_in_metro_areas, data = hate_crimes_minusDC)
summary(metro_SPLC_noDC)
```

### 4) Share of the population with a high school degree: (R-squared = 0.09545)

```{r, warning = FALSE}
high_school_SPLC <-lm(hate_crimes_per_100k_splc ~ share_population_with_high_school_degree, data = hate_crimes)
summary(high_school_SPLC)

ggplot(data = hate_crimes, aes(x = share_population_with_high_school_degree, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.1776)

```{r}
high_school_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ share_population_with_high_school_degree, data = hate_crimes_minusDC)
summary(high_school_SPLC_noDC)
```

### 5) Share of non-citizens: (R-squared = 0.0593)

```{r, warning = FALSE}
noncitizens_SPLC <-lm(hate_crimes_per_100k_splc ~ share_non_citizen, data = hate_crimes)
summary(noncitizens_SPLC)

ggplot(data = hate_crimes, aes(x = share_non_citizen, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.005073)

```{r}
noncitizens_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ share_non_citizen, data = hate_crimes_minusDC)
summary(noncitizens_SPLC_noDC)
```


### 6) Share of white poverty: (R-squared = 0.08189)

```{r, warning = FALSE}
white_poverty_SPLC <- lm(hate_crimes_per_100k_splc ~ share_white_poverty, data = hate_crimes)
summary(white_poverty_SPLC)

ggplot(data = hate_crimes, aes(x = share_white_poverty, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.009772)

```{r}
white_poverty_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ share_white_poverty, data = hate_crimes_minusDC)
summary(white_poverty_SPLC_noDC)
```

### 7) Gini index (measure of inequality): (R-squared = 0.1085)

```{r, warning = FALSE}
gini_SPLC <- lm(hate_crimes_per_100k_splc ~ gini_index, data = hate_crimes)
summary(gini_SPLC)

ggplot(data = hate_crimes, aes(x = gini_index, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.01286)

```{r}
gini_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ gini_index, data = hate_crimes_minusDC)
summary(gini_SPLC_noDC)
```

### 8) Share of residents who are non-white: (R-squared = 0.002262)

```{r, warning = FALSE}
nonwhite_SPLC <- lm(hate_crimes_per_100k_splc ~ share_non_white, data = hate_crimes)
summary(nonwhite_SPLC)

ggplot(data = hate_crimes, aes(x = gini_index, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.07141)

```{r}
nonwhite_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ share_non_white, data = hate_crimes_minusDC)
summary(nonwhite_SPLC_noDC)
```


### 9) Share of voters who voted for Trump in 2016: (R-squared = 0.4317)

```{r, warning = FALSE}
trump_SPLC <- lm(hate_crimes_per_100k_splc ~ share_voters_voted_trump, data = hate_crimes)
summary(trump_SPLC)

ggplot(data = hate_crimes, aes(x = share_voters_voted_trump, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Minus D.C. (R-squared = 0.1745)

```{r}
trump_SPLC_noDC <- lm(hate_crimes_per_100k_splc ~ share_voters_voted_trump, data = hate_crimes_minusDC)
summary(trump_SPLC_noDC)
```

For the SPLC data, the explanatory variables with the three highest correlations were: 

1) `share_voters_voted_trump` (R-squared = 0.4317, r = -0.657)

2) `median_household_income` (R-squared = 0.123, r = 0.351)

3) `gini_index` (R-squared = 0.1085, r = 0.329)


Once again, of the two variables that we predicted would be the most correlated, only one appears in this list. Additionally, the direction of the relationship does not match what was expected. `share_voters_voted_trump` has a negative relationship with hate crime rate here.

When the Washington, D.C. outlier was removed, the three most correlated variables were:

1) `share_population_with_high_school_degree` (R-squared = 0.1776, r = 0.421)

2) `share_voters_voted_trump` (R-squared = 0.1745, r = -0.273)

3) `median_household_income` (R-squared = 0.08232, r = 0.287)

In this situation, both variables we predicted would be the top two are in the expected positions. However, once again, both of their directions are opposite from what was predicted.

Just as with the FBI data, the two lists here are different. Median household income and share voters voted Trump are included on both lists, while gini index and share of the population with a high school degree are not.


### Finding the top model

In order to fit the best model, backward-elimination was used. The best fit model was calculated while including the data for Washington, D.C. Below is the model with all variables included.

```{r, warning = FALSE}

SPLC_model_all <- lm(hate_crimes_per_100k_splc ~ median_household_income + share_unemployed_seasonal + share_population_in_metro_areas + share_population_with_high_school_degree + share_non_citizen + share_white_poverty + gini_index + share_non_white + share_voters_voted_trump, data = hate_crimes)
summary(SPLC_model_all)

```

The best model for the SPLC data had more independent variables than the FBI data. The four variables that were statistically significant were `median_household_income` (p-value = 0.04349), `share_population_with_high_school_degree` (p-value = 0.00112), `gini_index` (p-value = 0.02856), and `share_voters_voted_trump` (p-value = 0.00180).

```{r, warning = FALSE}

SPLC_model_best <- lm(hate_crimes_per_100k_splc ~ median_household_income + share_population_with_high_school_degree + gini_index + share_voters_voted_trump, data = hate_crimes)
summary(SPLC_model_best)

```

The equation for this model is seen below:

hate_crimes_per_100k_splc = (-1.015e-5 x median_household_income) + (3.597 x share_population_with_high_school_degree) + (4.595 x gini_index) - (1.310 x share_voters_voted_trump) - 4.581


## How much can these models be trusted? (Conditions)

In order to determine whether it is valid to use these models for inference, it is important to check several conditions. We checked these for both the FBI model and the SPLC model.


### FBI

**1) Linearity:** Does each explanatory variable have a linear relationship with the response variable?

`gini_index`: This variable does not seem to have a linear relationship. There is a strange pattern of points. 

```{r}
ggplot(data = hate_crimes, aes(x = gini_index, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

`share_population_with_high_school_degree`: For this variable, it is difficult to tell whether there is actually a linear relationship, since there is not a steep line. However, there does not appear to be a relationship of any other type, so it might be reasonable to assume this is linear.

```{r}
ggplot(data = hate_crimes, aes(x = share_population_with_high_school_degree, y = avg_hatecrimes_per_100k_fbi)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

Due to `gini_index` having an irregular pattern, this condition is not met.

**2) Independence**: Are all of the observations independent of one another?

We can NOT assume that this condition is met. It is highly likely that since these observations are entire states, and states can be linked, that the cases are dependent on one another.

**3) Normality of residuals**: Are all of the residuals normally distributed?

```{r}
qplot(sample = .resid, data = FBI_model_best, stat = "qq") + ggtitle ("QQ Plot - FBI")
```

There is not a perfect normal shape here. The center section looks relatively normal, but the two extremes follow a much less normal shape. The residuals are not entirely normally distributed, but they do not appear far enough from the normal line to declare this condition not met. However, this quantile-quantile plot does provide enough evidence of non-normality to raise doubts about the model.

**4) Equal Variance of Residuals**: Are the points relatively equidistant from the least squares line at all x-values?

```{r}
qplot(x = .fitted, y = .resid, data = FBI_model_best) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals") + ggtitle ("Residuals vs. Fitted Values - FBI Model")
```

The points do not appear equidistant from the least squares line across all x-values. This condition does not appear to be entirely met.

**Collinearity?**

```{r}
hate_crimes %>%
  select("share_population_with_high_school_degree", "gini_index") %>%
  ggpairs()
```


Due to inability to clearly meet conditions 1, 3, and 4, this model should be known to not be entirely accurate. However, there does not appear to be any other noticable shape to the data that could replace a linear model, so this model type is likely the best predictor available.

### SPLC

**1) Linearity:** Does each explanatory variable have a linear relationship with the response variable?

`gini_index`: 

```{r}
ggplot(data = hate_crimes, aes(x = gini_index, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

There does not appear to be a clear linear relationship here. It is likely that the outlier represinting Washington D.C. is influencing the slope of the regression line.

`share_population_with_high_school_degree`: 

```{r}
ggplot(data = hate_crimes, aes(x = share_population_with_high_school_degree, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

There appears to be a slight positive slope and linear relationship here. This variable appears to meet the condition.

`median_household_income`:

```{r}
ggplot(data = hate_crimes, aes(x = median_household_income, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

It is difficult to tell whether there is a linear relationship here. Once again the outlier appears to be influencing the slope. However, no shape appears to better fit these data points than a line.

`share_voters_voted_trump`:

```{r}
ggplot(data = hate_crimes, aes(x = share_voters_voted_trump, y = hate_crimes_per_100k_splc)) + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

These data points do appear to follow a linear pattern, so they meet this condition.


**2) Independence**: Are all of the observations independent of one another?

We can NOT assume that this condition is met. It is highly likely that since these observations are entire states, and states can be linked, that the cases are dependent on one another.

**3) Normality of residuals**: Are all of the residuals normally distributed?

```{r}
qplot(sample = .resid, data = SPLC_model_best, stat = "qq") + ggtitle("QQ Plot - SPLC")
```

The data points here do not appear to entirely follow a normal model. The quantile-quantile plot appears to show most points differing from their expected values, though only a little bit. This condition is not clearly met, but also not clearly not met.

**4) Equal Variance of Residuals**: Are the points relatively equidistant from the least squares line at all x-values?

```{r}
qplot(x = .fitted, y = .resid, data = SPLC_model_best) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  xlab("Fitted values") +
  ylab("Residuals") +
  ggtitle ("Residuals vs. Fitted Values - SPLC Model")
```

The residuals do not appear to have equal variance. Therefore this condition is not met.


**Collinearity?**

```{r}
hate_crimes %>%
  select("median_household_income", "share_population_with_high_school_degree", "gini_index", "share_voters_voted_trump") %>%
  ggpairs()
```

Due to the fact that condition 4 is clearly not met, several variables from condition 1 are not clearly linear, and condition 3 is also not entirely met, it is important to proceed with caution when concluding based on the calculated model.

In addition, most variables in this model appear to show some collinearity. This is logical, since demographic characteristics are often related to one another. Due to this observation, it is important to note that the coefficients for many of the variables are being affected due to the fact that collinear variables are also included in the model. 


# Conclusion


Based on the calculations and results, the two research questions can be answered as follows:

**1) Are Trump voter rate and share of population with a high school degree the two variables most strongly associated with hate crimes? Do the directions of their relationships match what is expected?**

Using both the FBI and SPLC data (including the data for Washington D.C.), `share_voters_voted_trump` is one of the most correlated variables. `share_population_with_high_school_degree` is not. When checking the correlation direction for `share_voters_voted_trump`, we found that it was NOT what was expected. We predicted a positive correlation, but found a negative one. Further research would likely be required to determine why that was the case.

Since Washington D.C. is a clear outlier, as it is a city while all other observations represent entire states, we removed it from the dataset and completed the same calculations once again to determine if it was having an effect on our results. In this situation, one of our two predicted variables, `share_population_with_high_school_degree` was one of the most correlated variables with the FBI hate crime rate. For the SPLC hate crime rate, our prediction was correct: `share_voters_voted_trump` and `share_population_with_high_school_degree` were the two variables with the strongest corelations. However, neither variable's correlation direction matched what had been predicted. `share_voters_voted_trump` once again had a negative correlation, while `share_population_with_high_school_degree` had a positive correlation when a negative one was expected.

It is important to note that, for both of these results, correlation is not the same as causation. Although several of the explanatory variables had relatively large correlations, that does not mean that they directly caused more hate crimes to happen. For instance, with the negative correlation between `share_voters_voted_trump` and hate crime rate, we could not definitively say that removing a significant portion of Trump voters in any given area would increase the hate crime rate. There is no cause and effect implied by this result. 

In order to determine cause and effect, a randomized experiment would have to be completed. However, this would be illegal to conduct since it would involve committing hate crimes. We must be satisfied with correlation in this situation.


**2) What model can we create to best predict hate crime rate?**

The two models we created, using backward-selection with p-values were:


*`avg_hatecrimes_per_100k_fbi` = (31.287 x `share_population_with_high_school_degree`) + (64.324 x `gini_index`) - 54.017*

*`hate_crimes_per_100k_splc` = (-1.015e-5 x `median_household_income`) + (3.597 x `share_population_with_high_school_degree`) + (4.595 x `gini_index)` - (1.310 x `share_voters_voted_trump`) - 4.581*

In both models, `gini_index` and `share_population_with_high_school_degree` were significant variables. The FBI model only had two significant variables, which is very few considering the number of explanatory variables, and the SPLC model only had four. The SPLC model appears to also be a better predictor of hate crime rate, with an adjusted R-squared value of 0.5248 compared with the FBI value of 0.408. It is important to note, however, that neither model is particularly strong. While the SPLC model is better than the FBI, 0.5248 is still a relatively small R-squared.

Additionally, neither model meets all of the conditions for inference. Constant variability and normality of residuals are in question for both models, as well as independence. Since the states are not independent observations, values in some states may be affecting others. Additionally, there is a significant amount of collinearity between the variables included in the models. Since many economic and demographic statistics are interrelated, including multiple in the model affects the coefficients of all of them. However, since almost all of the variables are strongly correlated with one another, there is no way to remove variables from the model until none of them are associated.

The difficulty with studying, and forming predictive models, based on hate crime data, is that the data we have available is both connected and inaccurate to reality. The reported hate crime rates of both the FBI and SPLC are likely significantly lower than the true hate crime rates, and state crime rates themselves vary due to hate crime laws not being uniform across states. 

In order to form a more accurate model, we would have to look at a more complete record of hate crime data. To obtain this, hate crime laws would need to be made uniform across all states, and states would need to report the majority of hate crimes that happened. This would help solve some problems with the model, but of course could not solve the problem of independence--a problem which appears impossible to fix. One solution that could help improve this could be to possibly create cases that represent a smaller area, perhaps each case being a county rather than a state. Unfortunately, this is not data available to the public right now.

The model as it exists, could certainly serve to help average individuals belonging to minority groups that can be targeted in hate crimes determine areas they might want to avoid living in. While it is not perfect, it is better than no model at all, and we do not currently have data available to create a model that is much better than this one.

However, there are certainly more questions to be asked so that we can learn more. Specifically, what is causing the opposite associations from expectation that we observed with `share_voters_voted_trump` and `share_population_with_high_school_degree`? How would the hate crime model change if we considered smaller areas as our cases? 

Additionally, further breakdown of hate crime rates by type of crime--anti-LGBT, anti-Muslim, anti-Black, etc.--could allow for more models that would help individuals of certain groups. For instance, if we had data on the anti-LGBT hate crime rate of an area, and the demographic statistics (which might also then include information about the LGBT population), we could create a model for LGBT individuals which specifically predicts the level of hate crime against them, rather than just hate crimes in general. 


# References

Fivethirtyeight. “Fivethirtyeight/Data.” *GitHub*, 
  github.comfivethirtyeight/data/blob/master/hate-crimes/hate_crimes.csv.

Fivethirtyeight. "Education, Not Income, Predicted Who Would Vote for Trump." 
  *FiveThirtyEight*, 22 Nov. 2016, http://fivethirtyeight.com/features/
  education-not-income-predicted-who-would-vote-for-trump/.

Federal Bureau of Investigation (FBI). "Hate Crimes." *FBI.gov*, 
  https://www.fbi.gov/investigate/civil-rights/hate-crimes.

Southern Poverty Law Center (SPLC). “Report A Hate Incident.” *splcenter.org*, 
  https://www.splcenter.org/reporthate

Schwencke, Ken. “Why America Fails at Gathering Hate Crime Statistics.” *ProPublica*, 4 Dec. 
  2017. https://www.propublica.org/article/why-america-fails-at-gathering-hate-crime-statistics


















